---
title: "Data processing"
author: "Viola Wiegand"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
    toc: yes
---
```{r set-options, echo=FALSE, cache=FALSE}
options(width = 1000)
```

# Creating corp\_text objects

In order to run the CorporaCoCo comparison on the semantically tagged files, they have to be imported into R as `corp-text` objects. This was done using the following R code (based on code written by Anthony Hennessey for Viola Wiegand's thesis data). The code only works with the development version 1.1 of CorporaCoCo (i.e. the version that understands "complex types" and corp_text objects).

```{r create RDS files}
library(CorporaCoCo) # development version of 1.1
library(data.table)
library(stringi)

write_corp_text <- function(indir, outdir) {
    infiles <- list.files(indir, pattern = "\\.sem")

    for(file in infiles) {
        f <- readLines(file.path(indir, file))
        # this is probably fixed width but not documented, so separate on white space
        # the last column is tagged separated by white space 
        f <- stri_match(f, regex = "^(\\S+)\\s+(\\S+)\\s+(\\S+)\\s+(\\S+)\\s+(.*?)\\s*$")  # erronious white space at the end

        ff <- as.data.table(f[f[, 4] != "-----", 2:6], stringsAsFactors = FALSE)
        names(ff) <- c("col1", "col2", "POS", "token", "USAS")

        # undo HTML encoding we did for POS tagger
        html_entities <- list(
            from = c("&pound;", "&amp;", "&lt;", "&gt;", "&lsqb;", "&rsqb;", "&bquo;"),
            to = c("£", "&", "<", ">", "[", "]", "`")
    )
    tmp <- html_entities$to[match(ff$token, html_entities$from )]
    ff$token <- ifelse(is.na(tmp), ff$token, tmp)
        
        # we want cooccurrence barriers each time col1 changes
        idxs <- data.frame(starts = data.table:::uniqlist(ff[, 1]))
        idxs$ends <- shift(idxs$starts, -1) - 1
        idxs$ends[length(idxs$ends)] <- nrow(ff)
        n <- 5 # assuming span =< 5
        bar <- data.table("col1" = rep(NA, n), "col2" = rep(NA, n),  "POS" = rep(NA, n), "token" =  rep("_b_", n), "USAS" = rep(NA, n))
        fff <- do.call(rbind, apply(idxs, 1, function(x) {rbind(ff[x[1]:x[2],], bar)}))

        # attempt to regain some white space integrity
        fff$text <- ifelse(shift(fff$token, 1, type = "lead") %in% c(".", ",", ":", ";", "]", ")") | fff$token %in% c("(", "["),
            fff$token,
            paste0(fff$token, " ")
        )

        fff$nchar_token <- nchar(fff$token, type = "chars")
        fff$nchar_text <- nchar(fff$text, type = "chars")

        fff$start <- cumsum(c(1, fff$nchar_text))[-nrow(fff)]
        fff$end <- fff$start + fff$nchar_token - 1

        wanted <- fff$USAS != "" | is.na(fff$USAS)
        obj <- corp_text(
            text = paste(fff$text, collapse = ""),
            tokens = data.frame(
                type = fff$USAS[wanted],
                start = as.integer(fff$start[wanted]),
                end = as.integer(fff$end[wanted]),
                stringsAsFactors = FALSE
            )
        )

        outfile <- sub("\\.sem", ".sem.rds", file)
        saveRDS(obj, file = file.path(outdir, outfile))
    }
}

# DNov
## DNov quotes
indir <- "./api-output-sem-tagged/DNov/quote"
outdir <- "./api-output-sem-tagged/DNov/quote_corp_text_objects"
write_corp_text(indir, outdir)

## DNov non-quotes
indir <- "./api-output-sem-tagged/DNov/nonquote"
outdir <- "./api-output-sem-tagged/DNov/nonquote_corp_text_objects"
write_corp_text(indir, outdir)

## DNov long suspensions
indir <- "./api-output-sem-tagged/DNov/longsus"
outdir <- "./api-output-sem-tagged/DNov/longsus_corp_text_objects"
write_corp_text(indir, outdir)

## DNov short suspensions
indir <- "./api-output-sem-tagged/DNov/shortsus"
outdir <- "./api-output-sem-tagged/DNov/shortsus_corp_text_objects"
write_corp_text(indir, outdir)

# 19C
## 19C quotes
indir <- "./api-output-sem-tagged/19C/quote"
outdir <- "./api-output-sem-tagged/19C/quote_corp_text_objects"
write_corp_text(indir, outdir)

## 19C non-quotes
indir <- "./api-output-sem-tagged/19C/nonquote"
outdir <- "./api-output-sem-tagged/19C/nonquote_corp_text_objects"
write_corp_text(indir, outdir)

## 19C long suspensions
indir <- "./api-output-sem-tagged/19C/longsus"
outdir <- "./api-output-sem-tagged/19C/longsus_corp_text_objects"
write_corp_text(indir, outdir)

## 19C short suspensions
indir <- "./api-output-sem-tagged/19C/shortsus"
outdir <- "./api-output-sem-tagged/19C/shortsus_corp_text_objects"
write_corp_text(indir, outdir)

# ChiLit
## ChiLit quotes
indir <- "./api-output-sem-tagged/ChiLit/quote"
outdir <- "./api-output-sem-tagged/ChiLit/quote_corp_text_objects"
write_corp_text(indir, outdir)

## ChiLit non-quotes
indir <- "./api-output-sem-tagged/ChiLit/nonquote"
outdir <- "./api-output-sem-tagged/ChiLit/nonquote_corp_text_objects"
write_corp_text(indir, outdir)

## ChiLit long suspensions
indir <- "./api-output-sem-tagged/ChiLit/longsus"
outdir <- "./api-output-sem-tagged/ChiLit/longsus_corp_text_objects"
write_corp_text(indir, outdir)

## ChiLit short suspensions
indir <- "./api-output-sem-tagged/ChiLit/shortsus"
outdir <- "./api-output-sem-tagged/ChiLit/shortsus_corp_text_objects"
write_corp_text(indir, outdir)

# AAW
## AAW quotes
indir <- "./api-output-sem-tagged/AAW/quote"
outdir <- "./api-output-sem-tagged/AAW/quote_corp_text_objects"
write_corp_text(indir, outdir)

## AAW non-quotes
indir <- "./api-output-sem-tagged/AAW/nonquote"
outdir <- "./api-output-sem-tagged/AAW/nonquote_corp_text_objects"
write_corp_text(indir, outdir)

## AAW long suspensions
indir <- "./api-output-sem-tagged/AAW/longsus"
outdir <- "./api-output-sem-tagged/AAW/longsus_corp_text_objects"
write_corp_text(indir, outdir)

## AAW short suspensions
indir <- "./api-output-sem-tagged/AAW/shortsus"
outdir <- "./api-output-sem-tagged/AAW/shortsus_corp_text_objects"
write_corp_text(indir, outdir)
``` 